Evolution of the AGN UV luminosity function from redshift 7.5

### 2018-04-04 Paper/paper.tex
Update text

We now consider the implications of our AGN luminosity function models
for \HeII\ reionization.  We use the emissivity model developed in
Section~\ref{sec:e912} for this purpose.  The evolution of the volume-
averaged fraction $Q_\nHeIII$ of \HeIII\ is given by
\citep{2012ApJ...746..125H}
\begin{equation}
{dQ_\nHeIII\over dt}={\dot n_{\rm ion,4}\over \langle n_\nHe \rangle} -{Q_\nHeIII\over
\langle t_{\rm rec,He}\rangle},
\label{eqn:qHedot}
\end{equation}
where $\dot n_{\rm ion,4}$ is the number density of photons with
energy 4~Ry and above produced per unit time, $n_\nHe$ is the Helium
number density, and $\langle t_{\rm rec,He}\rangle$ is an average
recombination time scale for \HeIII.  We obtain the evolution of $\dot
n_{\rm ion,4}$ for AGN with $M_\mathrm{1450}<-21$ and
$M_\mathrm{1450}<-18$ by integrating the contribution of double power
law luminosity function obtained in distinct redshifts bins and
fitting \ref{eqn:e912fit} to the resultant \HeII-ionizing
emissivities.  We extrapolate the resultant fit to $z>7$, although the
contribution from this to \HeII-reionization is small.  The SEDs from
Equation~(\ref{eqn:sed}) and (\ref{eqn:sed_faint}) are assumed for
bright and faint AGN, respectively.  The recombination time scale in
Equation~(\ref{eqn:qHedot}) is given by
\begin{equation}
\langle t_{\rm rec}\rangle=[(1+2\chi) \langle n_\nH\rangle \alpha_B\,C]^{-1},
\label{eq:trec}
\end{equation}
where $n_\nH$ is the volume-averaged density of hydrogen, $\alpha_B$
is the case-B \HeIII\ recombination rate, and $\chi=0.083$ is the
cosmic number fraction of Helium.  We assume that the helium clumping
factor $C$ is given by \citep{2015ApJ...813L...8M}
\begin{equation}
  C = 2.8\left(\frac{1+z}{6}\right)^{-1.1}.
\end{equation}
We use the fitting function provided by \citet{1997MNRAS.292...27H}
for the recombination coefficient.  Note that
Equation~\ref{eqn:qHedot} neglects the \HeII\ Lyman-limit systems that
play an increasingly important role towards the end of Helium
reionization \citep{2009MNRAS.395..736B, 2017ApJ...851...50M}.  In the
absence of these self-shielded systems of high-density \HeII,
$Q_\nHeIII$ can continue to increase beyond unity.  We set it to one
when this happens.  This can be corrected by accounting for the
\HeII\ column density distribution and filtering the \HeII-ionizing
radiation background through it.  Unfortunately, the \HeII\ column
density distribution is itself uncertain.  We therefore consider our
treatment of $Q_\nHeIII$ accurate enough for the purpose of this
work.

The resultant reionization histories of \HeII\ are shown in
Figure~\ref{fig:qhe} for $M_\mathrm{1450}<-21$ and
$M_\mathrm{1450}<-18$ in red and blue, respectively.  We find that
Helium reionization is complete at $z\sim 3.2$ and $3.1$ in these
models, respectively, with a one-sigma uncertainty of about $\Delta
z\sim 0.2$.  This is likely consistent with the \HeII\ \lya effective
optical depth measurements of \citet{2016ApJ...825..144W}, who find
that the effective optical depth rises at $z>2.7$, but the incidence
of low optical depth sightlines beyond this redshift implies that the
substantial volume of \HeII\ was ionized already at $z=3.4$.
\HeII\ reionization begins at $z\gtrsim 5$ in our models.  This could
explain the transmissive sightlines reported by
\citet{2016ApJ...825..144W}, although a comparison with their data
requires a more detailed modelling of the IGM.

In Figure~\ref{fig:qhe}, we also show the reionization histories from
other models of \HeII\ reionization in the literature.  Helium
reionization is delayed in the model of \citet{2012ApJ...746..125H}
due to their reduced emissivity at $z\sim 3$.  In the
\citet{2015ApJ...813L...8M} model, a significant \HeII\ fraction is
ionized already at $z=6$ due to the very high emissivity assumed in
this model.  In the radiation-hydrodynamic simulations by
\citet{2016ApJ...828...90L}, the AGN population is calibrated to the
SDSS and COSMOS data at $z>3.5$ \citep{2012ApJ...755..169M,
  2013ApJ...768..105M}, combined with a model for quasar light curves.
The fiducial \HeII\ reionization history obtained by these authors is
shown in Figure~\ref{fig:qhe} along with an estimate of the
uncertainty.  Reionization ends considerably later in this model
($z\sim 2.6$).  Our model curves are closest in agreement to the
simulations of \citet{2009ApJ...694..842M} and the analytical model of
\citet{2018arXiv180104931P}.  Still, reionization completes earlier in
our model, partly due to the absence of \HeII\ Lyman-limit systems.
Our redshift of reionization agrees more closely with that in the
simulations presented by \citet{2014MNRAS.445.4186C}.

>>>>

We now consider the implications of our AGN luminosity function models
for \HeII\ reionization.  We use the emissivity model developed in
Section~\ref{sec:e912} for this purpose.  The evolution of the
volume-averaged fraction $Q_\nHeIII$ of \HeIII\ is given by
\citep{2012ApJ...746..125H}
\begin{equation}
{dQ_\nHeIII\over dt}={\dot n_{\rm ion,4}\over \langle n_\nHe \rangle} -{Q_\nHeIII\over
\langle t_{\rm rec,He}\rangle},
\label{eqn:qHedot}
\end{equation}
where $\dot n_{\rm ion,4}$ is the number density of photons with
energy 4~Ry and above produced per unit time, $n_\nHe$ is the Helium
number density, and $\langle t_{\rm rec,He}\rangle$ is an average
recombination time scale for \HeIII.  We obtain the evolution of $\dot
n_{\rm ion,4}$ for AGN with $M_\mathrm{1450}<-21$ and
$M_\mathrm{1450}<-18$ by integrating the contribution of double power
law luminosity function obtained in distinct redshifts bins and
fitting Equation~(\ref{eqn:e912fit}) to the resultant \HeII-ionizing
emissivities.  We extrapolate the resultant fit to $z>7$, although the
contribution from this extrapolated emissivity to \HeII-reionization
is small.  As before, we assume the SEDs given by
Equations~(\ref{eqn:sed}) and (\ref{eqn:sed_faint}) for bright and
faint AGN, respectively.  The recombination time scale in
Equation~(\ref{eqn:qHedot}) is given by
\begin{equation}
\langle t_{\rm rec}\rangle=[(1+2\chi) \langle n_\nH\rangle \alpha_B\,C]^{-1},
\label{eq:trec}
\end{equation}
where $n_\nH$ is the average physical number density of hydrogen,
$\alpha_B$ is the case-B \HeIII\ recombination rate, and $\chi=0.083$
is the cosmic number fraction of Helium.  We assume that the helium
clumping factor $C$ is given by \citep{2015ApJ...813L...8M}
\begin{equation}
  C = 2.9\left(\frac{1+z}{6}\right)^{-1.1}.
\end{equation}
We use the fitting function provided by \citet{1997MNRAS.292...27H}
for the recombination coefficient.

Note that Equation~(\ref{eqn:qHedot}) neglects the \HeII\ Lyman-limit
systems that play an increasingly important role towards the end of
Helium reionization \citep{2009MNRAS.395..736B, 2017ApJ...851...50M}.
In the absence of these self-shielded systems of high-density \HeII,
$Q_\nHeIII$ can continue to increase beyond unity.  We set it to one
when this happens.  This can be corrected by accounting for the
\HeII\ column density distribution and filtering the \HeII-ionizing
radiation background through it.  Unfortunately, the \HeII\ column
density distribution is itself uncertain.  We therefore consider our
treatment of $Q_\nHeIII$ accurate enough for the purpose of this work.

The resultant reionization histories of \HeII\ are shown in
Figure~\ref{fig:qhe} for $M_\mathrm{1450}<-18$ and
$M_\mathrm{1450}<-21$ in red and blue, respectively.  We find that
Helium reionization is complete, respectively, at $z\sim 3.2$ and
$3.1$ in these models, with a one-sigma uncertainty of about $\Delta
z\sim 0.2$.  Observational constraints on the \HeII\ reionization
history are derived from measurements of the effective \HeII\ \lya
optical depth \citep{2001Sci...293.1112K, 2004ApJ...600..570S,
  2006A&A...455...91F, 2010ApJ...722.1312S, 2011ApJ...733L..24W,
  2016ApJ...825..144W}.  While we do not compute the effective
\HeII\ \lya optical depth in our model, the reionization histories
shown in Figure~\ref{fig:qhe} may likely be consistent with the
optical depth measurements of \citet{2016ApJ...825..144W}, who find
that the effective optical depth rises at $z>2.7$, but low optical
depth sightlines continue to be seen beyond this redshift.  The
incidence of these low opacity sightlines implies that the substantial
volume of \HeII\ was ionized already at $z=3.4$.  \HeII\ reionization
begins at $z\gtrsim 5$ in our models.  When we consider AGN with
$M_\mathrm{1450}<-18$, the ionization fraction $Q_\nHeIII$ has a
one-sigma spread of 0.8--1 at $z=3.4$.

In Figure~\ref{fig:qhe}, we also show the reionization histories from
other models of \HeII\ reionization in the literature.  Our model
curves are closest in agreement to the simulations of
\citet{2009ApJ...694..842M} and the analytical model of
\citet{2018arXiv180104931P}.  \citet{2018arXiv180104931P} model the
\HeII-ionizing emissivity of AGN by fitting the function form in
Equation~(\ref{eqn:e912fit}) to the emissivities reported by
\citet{2009A&A...507..781S}, \citet{2007A&A...472..443B}, and
\citet{2012ApJ...755..169M}.  The resultant emissivity in the model of
\citet{2018arXiv180104931P} is similar in shape but higher in
amplitude than the emissivity in the \citet{2012ApJ...746..125H}
model.  The \HeII-ionizing emissivity of \citet{2018arXiv180104931P}
is closer to our inference for $M_\mathrm{1450}<-21$ than the model of
\citet{2012ApJ...746..125H}.  (However, note that both these models
suffer from inhomogeneous luminosity function integration limits, as
the integration limit is set to a fixed fraction of the break
luminosity (cf.\ discussion in Section~\ref{sec:e912}.)  A spectral
index of $\alpha_\mathrm{UV}=-1.7$ is assumed.
\citet{2018arXiv180104931P} compute the evolution of $Q_\nHeIII$ by
evolving the \HeIII\ ionization fraction in a gas cloud with the mean
cosmic baryonic density exposed to the average \HeII-ionizing flux.
The ionizing flux is obtained by filtering the ionizing emissivity
through an inhomogeneous IGM in which the column density distribution
of \HeIII\ is derived from that of \HI\ in a manner similar to
\citet{2012ApJ...746..125H}.  As a result, this model incorporates the
effect of \HeII\ Lyman-limit systems, and the $Q_\nHeIII$ evolution in
this model deviates from our predictions towards the end of
reionization, as seen in Figure~\ref{fig:qhe}.  Comparing the
emissivity model in the cosmological radiative transfer simulations of
\citet{2009ApJ...694..842M} with our models is less straightforward.
Quasars are modelled in the simulations of \citet{2009ApJ...694..842M}
by relating quasar luminosities and lifetimes to halo masses following
the galaxy-merger-induced quasar activity model of
\citet{2005ApJ...630..705H}.  The resultant bolometric quasar
luminosity functions are somewhat shallower than those of
\citet{2007ApJ...654..731H}, but with a higher amplitude.  A spectral
index of $\alpha_\mathrm{UV}=-1.6$ is assumed.  Figure~\ref{fig:qhe}
shows that the resultant reionization history in these simulations
agrees with our model for $M_\mathrm{1450}<-21$, except towards the
end of reionization, where, as before, \HeII\ Lyman-limit systems
become important.

Helium reionization is delayed in the model of
\citet{2012ApJ...746..125H} due to their reduced emissivity at $z\sim
3$.  In the \citet{2015ApJ...813L...8M} model, a significant
\HeII\ fraction is ionized already at $z=6$ due to the very high
emissivity assumed in this model.  In the radiation-hydrodynamic
simulations by \citet{2016ApJ...828...90L}, the AGN population is
calibrated to the SDSS and COSMOS data at $z>3.5$
\citep{2012ApJ...755..169M, 2013ApJ...768..105M}, combined with a
model for quasar light curves.  The fiducial \HeII\ reionization
history obtained by these authors is shown in Figure~\ref{fig:qhe}
along with an estimate of the uncertainty.  Reionization ends
considerably later in this model ($z\sim 2.6$).  Our redshift of
reionization agrees more closely with that in the simulations
presented by \citet{2014MNRAS.445.4186C}.
